---
title: "p8105_hw6_yy3307"
output: github_document
date: "2022-11-29"
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
```

## Problem 1


## Problem 2

First we need to import and clean the `homicide` dataset. Then I'm going to create a `city_state` variable as combination of `city` and `state`, with a binary variable indicating whether the homicide is solved. `victim_age` will be `mutate` as numeric expression. Dallas, TX; Phoenix, AZ; Kansas City, MO; Tulsa, AL are being omitted and we only focus our analysis on those for whom `victim_race` is `white` or `black`.

```{r, warning = FALSE}
homicide_raw = 
  read.csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    city_state = str_c(city, ", ", state),
    solve_binary = as.numeric(
      ifelse(disposition == "Closed by arrest", 1, 0)
    ),
    victim_age = as.numeric(victim_age)
  ) %>% 
  filter(
    city_state != "Dallas, TX", 
    city_state != "Phoenix, AZ", 
    city_state != "Kansas City, MO", 
    city_state != "Tulsa, AL"
  ) %>% 
  filter(victim_race %in% c("White", "Black"))
```

Then I'm going to use the `glm` function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors for city of Baltimore, MD. After applying `broom::tidy` to it, we need to calculate the estimate and confidence interval.

```{r}
baltimore_logistic =
  homicide_raw %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(solve_binary ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_upper = exp(estimate - 1.96 * std.error),
    CI_lower = exp(estimate - 1.96 * std.error)
  ) %>% 
  select(term, log_OR = estimate, OR, CI_lower, CI_upper) %>% 
  knitr::kable(digits = 3)

baltimore_logistic
```


## Problem 3

